---
title: "Reviewer Comments - Pipeline Comparison"
output:
  workflowr::wflow_html:
    toc: true
    toc_float: true
    theme: cosmo
    highlight: zenburn
    code_folding: hide
editor_options:
  chunk_output_type: console
---



```{r setup, message=FALSE, warning=TRUE}
library(ggplot2)
library(viridis)
library(patchwork)

library(readr)
library(dplyr)
library(tibble)
```



```{r, message=FALSE, warning=TRUE}
source("code/core_settings.R")
source("code/core_functions.R")
```



```{r}
# set parameters
fileID = "normal_subset_neuronal_glial_cleaned"
leiden_res = 'leiden_0_4'
leiden_res_annot = 'annot_leiden_0_4'

adata_path = paste0('data/forebrain_', fileID, '.h5ad')


py_run_string("import numpy as np")
py_run_string("import anndata as ad")

py_run_string("adata = ad.read_h5ad(r.adata_path)")



metadata_scanpy = as_tibble(py$adata$obs, rownames="cellID") %>%
                    rename(mt_frac = percent_mito) %>%
                    mutate(annot_scanpy = annot_leiden) %>%
                    select("cellID", "sampleID", "gate", "timepoint", "set",
                           "treatment", "condition", "annot_scanpy") %>%
                    mutate(barcode = purrr::map_chr(stringr::str_split(cellID, '-'), 1)) %>%
                    mutate(FACS_gate = gsub("SOX2_", "GFP", gate)) %>%
                    mutate(barcode_gate_timepoint = paste(barcode, FACS_gate, timepoint, sep="_"))


metadata_seurat = read_csv("data/external_metadata/metadata_neuronal_glial_seurat.csv") %>%
                  rename(annot_seurat = labels) %>%
                  rename(FACS_gate = gfp) %>%
                  rename(cellID_seurat = barcode) %>%
                  select(cellID_seurat, orig.ident, seurat_clusters, timepoint, sample,
                         FACS_gate, annot_seurat) %>%
                  mutate(barcode = purrr::map_chr(stringr::str_split(cellID_seurat, '_'), 2)) %>%
                  mutate(timepoint = gsub("E12", "E12.5", timepoint)) %>%
                  mutate(timepoint = gsub("E14", "E14.5", timepoint)) %>%
                  mutate(timepoint = gsub("E16", "E16.5", timepoint)) %>%
                  mutate(barcode_gate_timepoint = paste(barcode, FACS_gate, timepoint, sep="_"))


metadata = full_join(metadata_scanpy %>% select(barcode_gate_timepoint, annot_scanpy),
                     metadata_seurat %>% select(barcode_gate_timepoint, annot_seurat),
                     by="barcode_gate_timepoint") %>%
                     mutate(annot_scanpy = tidyr::replace_na(as.vector(annot_scanpy), "Non-overlapping") %>% as.factor()) %>%
                     mutate(annot_seurat = tidyr::replace_na(as.vector(annot_seurat), "Non-overlapping") %>% as.factor())
                     

levels_annot_scanpy = c("Embryonic RG",
                        "Juvenile RG & TAPs",
                        "Gliogenic precursors (aNSCs)",
                        "Gliogenic precursors (APCs)",
                        "Quiescent NSCs (dorsal)",
                        "Quiescent NSCs (ventral)",
                        "Early EmNBs [1]",
                        "Early EmNBs [2]",
                        "GE NBs [1]",
                        "GE NBs [2]",
                        "Hippocampal NBs [1]",
                        "Hippocampal NBs [2]",
                        "OPCs",
                        "PreM-ODCs",
                        "Oligodendrocytes [1]",
                        "Oligodendrocytes [2]",
                        "Ependymal cells",
                        "GABAergic INs",
                        "ImStNeurons",
                        "Non-overlapping")


levels_annot_seurat = c("Radial glia-1",
                        "Radial glia-2 & TAPs",
                        "Active NSCs",
                        "APCs",
                        "Quiescent NSCs-1",
                        "Quiescent NSCs-2",
                        "Thalamic INs",
                        "Cajal-Retzius cells",
                        "SVZ NBs",
                        "GE progenitors",
                        "CE progenitors",
                        "IPCs & Hippoc. NBs & GCs",
                        "OPCs",
                        "Oligodendrocytes-2",
                        "Oligodendrocytes-1",
                        "Ependymal cells",
                        "Inhibitory GABA INs",
                        "Imm. striatal Neurons",
                        "Non-overlapping")


levels_annot_seurat_final = c("Embryonic RG",
                              "Juvenile RG & TAPs",
                              "Gliogenic precursor-1",
                              "Gliogenic precursor-2",
                              "qNSCs-1",
                              "qNSCs-2",
                              "Early EmDieNBs",
                              "Early EmNBs / CRCs",
                              "GE neuroblast 1",
                              "GE neuroblast 2",
                              "Hippocampal NBs-1",
                              "Hippocampal NBs-2",
                              "OPCs",
                              "Oligodendrocytes-2",
                              "Oligodendrocytes-1",
                              "Ependymal cells",
                              "GABAergic INs",
                              "ImStNeurons",
                              "Non-overlapping")

names(levels_annot_seurat_final) = levels_annot_seurat


tbl_cluster_stats = metadata %>%
                  group_by(annot_scanpy, annot_seurat) %>%
                  dplyr::summarise(n_counts = n()) %>%
                  ungroup() %>%
                  tidyr::complete(annot_scanpy, annot_seurat, fill = list(n_counts = 0)) %>%
                  group_by(annot_scanpy) %>%
                  mutate(frac_norm_scanpy = n_counts/sum(n_counts)) %>%
                  mutate(norm_scanpy = sum(n_counts)) %>%
                  ungroup() %>%
                  mutate(annot_scanpy = factor(annot_scanpy, levels=levels_annot_scanpy)) %>%
                  mutate(annot_seurat_v0 = factor(annot_seurat, levels=levels_annot_seurat)) %>%
                  mutate(annot_seurat = as.vector(annot_seurat))


for (cluster_annot in levels_annot_seurat){
  tbl_cluster_stats = tbl_cluster_stats %>%
                        mutate(annot_seurat = if_else(annot_seurat == cluster_annot,
                               levels_annot_seurat_final[cluster_annot],
                               annot_seurat))
}


tbl_cluster_stats = tbl_cluster_stats %>%
                      mutate(annot_seurat = factor(annot_seurat, levels=levels_annot_seurat_final))
```

```{r}
plt_heatmap_stats = ggplot(tbl_cluster_stats, aes(x=annot_seurat, y=annot_scanpy, fill=n_counts)) +
                            geom_tile() +
                            scale_fill_viridis() +
                            theme_publication +
                            scale_x_discrete(expand = c(0,0),
                                             guide = guide_axis(angle=45)) +
                            scale_y_discrete(expand = c(0,0)) +
                            labs(x="Annotation [Seurat]", y="Annotation [Scanpy]") +
                            labs(fill="n cells") +
                            guides(fill = guide_colorbar(barwidth = 0.3,
                                                         barheight = 2))
                                                         
plt_heatmap_stats
```


```{r}
plt_heatmap_stats = ggplot(tbl_cluster_stats, aes(x=annot_seurat, y=annot_scanpy, fill=frac_norm_scanpy)) +
                            geom_tile() +
                            scale_fill_viridis(limits=c(0, 1), breaks=c(0, 0.5, 1)) +
                            theme_publication +
                            scale_x_discrete(expand = c(0,0),
                                             guide = guide_axis(angle=45)) +
                            scale_y_discrete(expand = c(0,0)) +
                            labs(x="Annotation [Seurat]", y="Annotation [Scanpy]") +
                            labs(fill="Fraction [Scanpy]") +
                            guides(fill = guide_colorbar(barwidth = 0.3,
                                                         barheight = 2))
                                                         
plt_heatmap_stats
```

## Statistics

```{r}
n_not_scanpy = tbl_cluster_stats %>%
                  filter(annot_seurat == "Non-overlapping") %>%
                  pull(n_counts) %>%
                  sum()

n_not_seurat = tbl_cluster_stats %>%
                  filter(annot_scanpy == "Non-overlapping") %>%
                  pull(n_counts) %>%
                  sum()
                  
frac_overlap_max = tbl_cluster_stats %>%
                      filter(annot_scanpy != "Non-overlapping") %>% # remove - not a cluster
                      group_by(annot_scanpy) %>%
                      summarise(frac_norm_scanpy_max = max(frac_norm_scanpy))

frac_overlap_max_median = median(frac_overlap_max %>% pull(frac_norm_scanpy_max))

n_scanpy = metadata %>%
              filter(annot_scanpy != "Non-overlapping") %>%
              nrow()

n_seurat = metadata %>%
              filter(annot_seurat != "Non-overlapping") %>%
              nrow()
```

```{r}
# n_not_scanpy
# n_not_seurat
# 
# frac_overlap_max_median
```

Number of cells in scanpy set: `r n_scanpy`

Number of cells in seurat set: `r n_seurat`

Number of cells not in scanpy set: `r n_not_scanpy` (fraction of seurat: `r paste0(round(n_not_scanpy/n_seurat*100, 2), "%")`)

Number of cells not in seurat set: `r n_not_seurat` (fraction of scanpy: `r paste0(round(n_not_seurat/n_scanpy*100, 2), "%")`)

Median max frac overlap between clusters (scanpy) (i.e. median fraction of the best matches): `r paste0(round(frac_overlap_max_median*100, 2), "%")`

In addition, the clusters can be clearly classified using known marker genes or are separated by distinct developmental timepoints.

# Markers

```{r}
marker_list = c("Sox2", "Nes", "Hes1", "Prom1", "Aldoc", "Slc1a3", "Vcam1", "Ntsr2",
                 "Nhlh2", "Nhlh1", "Ebf2", "Lhx9", "Reln", "Irx1", "Irx2", "Irx3",
                 "Mki67", "Dcx", "Dlx1", "Dlx2", "Dlx5", "Sp9", "Neurod2", "Tbr1",
                 "Olig1", "Olig2", "Pdgfra", "Plp1", "Mbp", "Foxj1", "Gad2", "Gata2",
                 "Gata3", "Rarb")

metadata = metadata_scanpy

tbl_expr = get_adata_raw_expr(marker_list)
leiden_res_annot = "annot_scanpy"
annot_order = levels_annot_scanpy

tbl_DE_expr = metadata %>%
                select(cellID, !!(leiden_res_annot)) %>%
                mutate(clusterID = get(leiden_res_annot)) %>%
                left_join(tbl_expr, by="cellID") %>%
                tidyr::pivot_longer(starts_with("expr_"), names_to="geneID", values_to="expr") %>%
                mutate(geneID = gsub("expr_", "", geneID)) %>%
                group_by(geneID, clusterID) %>%
                summarise(frac_expressed = sum(expr > 0)/length(expr),
                          expr_average = mean(expr)) %>%
                ungroup() %>%
                group_by(geneID) %>%
                mutate(expr_scaled = expr_average/max(expr_average)) %>%
                ungroup() %>%
                mutate(geneID = factor(geneID, levels = marker_list)) %>%
                mutate(clusterID = factor(clusterID, levels=rev(annot_order))) %>%
                filter(!is.na(clusterID))

# expr_Gene
plt_DE_dotplot = ggplot(tbl_DE_expr, aes(x=geneID, y=clusterID, size=frac_expressed, colour=expr_scaled)) +
                            geom_point(stroke=0) +
                            scale_size_continuous(range = c(0, 2.0), limits=c(0, 1), breaks = c(0.1, 0.5, 1.0)) +
                            theme_publication +
                            scale_colour_distiller(palette = 'RdYlBu', breaks=c(0, 0.5, 1), limits=c(0,1)) +
                            labs(x=NULL, y="Annotation [Scanpy]") +
                            labs(colour = "Expr (scaled)") +
                            guides(colour = guide_colorbar(barwidth = 0.3,
                                                         barheight = 1.5)) +
                            theme(axis.text.x = element_text(angle = 90, vjust=0.5, hjust=1)) +
                            labs(size = "Fraction expressing") +
                            theme(legend.spacing.y = unit(0.05, 'cm')) +
                            theme(legend.key.height = unit(0.2, "cm"))
                            # legend.key.width = unit(0.1, "cm"))


```


```{r, fig.asp=0.7}
plt_DE_dotplot_neuronal_glial = plt_DE_dotplot
plt_DE_dotplot_neuronal_glial
```






```{r}
# set parameters
fileID = "normal_subset_neuronal_glial_cleaned"
leiden_res = 'leiden_0_4'
leiden_res_annot = 'annot_leiden'

adata_path = paste0('data/forebrain_', fileID, '.h5ad')

 
celltype_colours = neuronal_glial_colours
colour_celltypes = neuronal_glial_colours
annot_order = names(neuronal_glial_colours)

py_run_string("import numpy as np")
py_run_string("import anndata as ad")

py_run_string("adata = ad.read_h5ad(r.adata_path)")
```


```{r, message=FALSE, warning=TRUE}
metadata_scanpy = as_tibble(py$adata$obs, rownames="cellID") %>%
                    rename(mt_frac = percent_mito) %>%
                    mutate(UMAP_dim1 = py$adata$obsm[['X_umap']][,1],
                           UMAP_dim2 = py$adata$obsm[['X_umap']][,2])

metadata_experiment = read_tsv("data/metadata_experiment.tsv")
results_cyclone = read_csv("data/cyclone/cyclone_merged.csv.gz", col_types="ccddd") %>%
                    mutate(sampleID = sapply(cellID, function(cell) strsplit(cell, split="-")[[1]][3])) %>%
                    left_join(metadata_experiment, by="sampleID") %>%
                    mutate(cellID = gsub("-1", "", cellID)) %>%
                    select(cellID, cyclone, cyclone_G1, cyclone_S, cyclone_G2M)

results_logreg = read_csv(paste0("data/logreg/logreg_Zeisel_2018_normal.csv.gz"))


# filter QC
metadata = left_join(metadata_scanpy, results_cyclone, by="cellID") %>%
              mutate(cyclone = factor(cyclone, levels=c("G1", "S", "G2M"))) %>%
              mutate(timepoint = factor(timepoint, levels=unique(timepoint)[order(unique(postnatal_day))])) %>%
              left_join(results_logreg, by="cellID") %>%
              mutate(leiden = get(leiden_res))
```


# Timepoints


```{r}
library(ggrastr)
                            
plt_UMAP_day = ggplot(slice(metadata, sample(1:n())), aes(x=UMAP_dim1, y=UMAP_dim2, colour=timepoint)) +
                            # geom_point(size=0.05, alpha=0.8, shape=16) +
                            geom_point_rast(size=0.05, alpha=0.8, shape=16, raster.dpi=300) +
                            # scale_colour_viridis(discrete=T) +
                            scale_color_manual(values=colour_timepoints) +
                            theme_publication +
                            labs(x="UMAP 1", y="UMAP 2") +
                            theme(axis.text.x=element_blank(),
                                  axis.ticks.x=element_blank(),
                                  axis.text.y=element_blank(),
                                  axis.ticks.y=element_blank()) +
                            labs(colour="Time point") +
                            guides(colour = guide_legend(override.aes = list(size=2),
                                                         keyheight=0.3,
                                                         default.unit="cm"))
```

```{r}
plt_UMAP_day
```



# Cluster annotation


```{r}
library(ggrepel)
library(ggthemes)
library(ggrastr)

plot_UMAP_leiden <- function(leiden_key){
  
  metadata = metadata %>%
              mutate(leiden = get(leiden_key))
              
  cl_cent = metadata %>%
                  select(UMAP_dim1, UMAP_dim2, leiden) %>%
                  group_by(leiden) %>%
                  summarize_all(mean)
  
  # randomise order for plot
  plt_UMAP_leiden = ggplot(slice(metadata, sample(1:n())), aes(x=UMAP_dim1, y=UMAP_dim2, colour=leiden)) +
                              # geom_point(size=0.2, alpha=0.8, shape=16) +
                              geom_point_rast(size=0.2, stroke=0, alpha=0.8, shape=16, raster.dpi=600) +
                              # geom_point_rast(size=0.2, alpha=0.8, shape=16, raster.dpi=600) +
                              geom_label_repel(aes(x=UMAP_dim1, y=UMAP_dim2, label=leiden),
                                      fill='white', colour='black', label.size=NA,
                                      alpha=0.9, label.padding = 0.1, point.padding = 0,
                                      data=cl_cent, size=1.78, show.legend=FALSE, inherit.aes=F) +
                              scale_colour_manual(values=celltype_colours) +
                              theme_publication +
                              labs(x="UMAP 1", y="UMAP 2") +
                              theme(axis.text.x=element_blank(),
                                    axis.ticks.x=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank()) +
                              # guides(color = guide_legend(override.aes = list(size=3))) +
                              # labs(colour="Leiden")
                              # labs(title=paste0(leiden_key)) +
                              theme(legend.position="none")

  return(plt_UMAP_leiden)
}
```

```{r}
plt_UMAP_annot_neuronal_glial = plot_UMAP_leiden("annot_leiden")
plt_UMAP_annot_neuronal_glial
```


## Plot panel

```{r, fig.asp=0.3}
plt_panel_heatmap_overlap_markers = plt_heatmap_stats + plt_DE_dotplot_neuronal_glial
plt_panel_heatmap_overlap_markers
```

```{r, fig.asp=0.3}
plt_UMAP_panel_scanpy = plt_UMAP_annot_neuronal_glial + plt_UMAP_day
plt_UMAP_panel_scanpy
```

```{r}
ggsave("figures/figure_comparison_scanpy_seurat/panel_scanpy_seurat_neuronal_glial.pdf", plot=plt_panel_heatmap_overlap_markers + theme(legend.margin=margin(0, 0, 0, -2, unit='mm')), height=8, width=22, units="cm")

ggsave("figures/figure_comparison_scanpy_seurat/UMAP_scanpy_neuronal_glial.pdf", plot=plt_UMAP_panel_scanpy + theme(legend.margin=margin(0, 0, 0, -2, unit='mm')), height=7, width=18.3, units="cm")
```
